---
title: "Plotting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(FIPS)
library(tidyverse)
library(lubridate)
```

## Introduction

In this vignette we briefly demonstrate how to use the FIPS
plotting function. Below we create sleeptimes and simulate an example 
sleep scenario to plot. This example includes initial sleep deprivation.

```{r}
# Simulation start date time (i.e., when you want first predictions to begin)
simulation.start = lubridate::ymd_hms('2018-05-01 11:00:00', tz = "Australia/Perth")
# Simulation end date time (i.e., when you want predictions to end)
simulation.end = lubridate::ymd_hms('2018-05-09 21:00:00', tz = "Australia/Perth")
# Sleep times, 5 hours per night, starting 23:00
example.sleeptimes <- tibble::tibble(
  sleep.start = seq(
    from = lubridate::ymd_hms('2018-05-03 02:00:00', tz = "Australia/Perth"), 
    to = lubridate::ymd_hms('2018-05-09 16:00:00', tz = "Australia/Perth"),
    by = '24 hours'),
  sleep.end = sleep.start + lubridate::dhours(3),
  sleep.id = rank(sleep.start))

simulated.dataframe = parse_sleeptimes(
  sleeptimes = example.sleeptimes,
  series.start = simulation.start,
  series.end = simulation.end,
  sleep.start.col = "sleep.start",
  sleep.end.col = "sleep.end",
  sleep.id.col = "sleep.id",
  roundvalue = 5)

unified.simulation.results = FIPS_simulate(
  FIPS_df = simulated.dataframe, # The FIPS_df
  modeltype = "unified",         # three process model
  pvec = unified_make_pvec()      # paramater vector
  )

```

FIPS contains a generic `plot` method for simulation results. By default this will
plot fatigue over time. Sleep is indicated by the grey rectangles, and the start
of each new day is indicated by the dashed lines.

```{r fig.width=7}
plot(unified.simulation.results)
```

Often it will be helpful to restrict the plotting between certain datetimes,
and plot additional data such as dynamic model parameters:

```{r fig.width=7}
plot(unified.simulation.results, from='2018-05-03 13:30:00',
                     to = "2018-05-07 18:30:00", plot_stat=c("fatigue", "s", "c", "l"))
```

Observed data can be plotted against the sleep predictions. Below demonstrates
by sampling some data points from the previously generated data frame,
then adding some random noise to make some 'pretend' observed data.


```{r fig.width=7}

pretend_ratings <- unified.simulation.results %>% filter(wake_status) 

pretend_ratings <- pretend_ratings[sample(1:length(pretend_ratings$datetime), 100),
                   c("datetime", "sim_hours", "fatigue")]

pretend_ratings$fatigue <- pretend_ratings$fatigue + rnorm(length
                                                                         (pretend_ratings$fatigue),
                                                                         0,
                                                                         0.7)

plot(unified.simulation.results, from='2018-05-03 13:30:00',
                     to = "2018-05-07 18:30:00", plot_stat=c("fatigue", "s", "c", "l"),
                     fatigue_CIs = FALSE, observed=pretend_ratings, observed_y="fatigue")

```


Below demonstrates an example plotting the three process model. Results are
similar except the output is alertness, rather than fatigue. This is
automatically handled for you if you are using the generic `plot` function.

```{r fig.width=7}
TPM.simulation.results = FIPS_simulate(
  FIPS_df = simulated.dataframe, # The FIPS_df
  modeltype = "TPM",         # three process model
  pvec = TPM_make_pvec()      # paramater vector
  )

plot(TPM.simulation.results)
```

It is important to note that the generic `plot()` function is simply calling `FIPS_plot`
under the hood. Users wishing full control may find the `FIPS_plot` function more intuitive to work with.

```{r, fig.width=7}
FIPS_plot(TPM.simulation.results, 
          plot_stat = c("alertness", "c", "u", "s")) +
  ggplot2::ggtitle("Three Process Model Simulation") +
  ggplot2::scale_color_brewer(palette="Set1")

```

